Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-07 WSSSPE2_report/WSSSPE2_review.tex
Section 4. More minor changes.

\begin{comment}
Lighting talks, breakout groups, working groups are all different. Lightning
talks were new in WSSSPE2. They were offered to all paper contributors, and some
other attendees, those who got travel awards. Breakout groups in WSSSPE1 were
just discussions about topics; in WSSSPE2, the breakout groups were supposed to
come up with plans for actions in specific areas. The idea of working groups is
that they will come out of breakout groups as a way for the groups to actually
do the actions over time, not at the workshops.
\end{comment{

\begin{comment}  
Email addresses of speakers:
gvwilson@software-carpentry.org, Colin Venters <C.Venters@hud.ac.uk>,
j.spencer@imperial.ac.uk, erinrobinson@esipfed.org, marpierc@iu.edu,
jwpeterson@gmail.com, abani@buffalo.edu, clenhardt@renci.org, dsk@uchicago.edu,
Samin.Ishtiaq@microsoft.com, james@howison.name, s.a.harris@leeds.ac.uk,
marcus.hanwell@kitware.com, thabermann@hdfgroup.org, rdowns@ciesin.columbia.edu,
cboettig@gmail.com, jakob.blomer@cern.ch, editor@ascl.net 
\end{comment}

@@

\item Colin Venters: The Nebuchadnezzar Effect: Dreaming of Sustainable Software
through Sustainable Software Architectures

Venters proposed that technically sustainable software is a composite, first-class, 
non-functional requirement (NFR) that is at a minimum a measure of a systems 
maintainability and extensibility but may also include other NFR such as efficiency 
(energy and cost), interoperability, portability, reusability, scalability, and usability. To 
achieve technically sustainable software Venters suggested that software 
architectures are fundamental as they are the primary carrier of system qualities 
(NFR), i.e., pre-system understanding, and influence how developers are able to 
understand, analyze, extend, test and maintain a software system, i.e., post-
deployment system understanding. In addition, Venters highlighted that the 
sustainability of software architectures also needs to be addressed to endure 
different types of change and evolution in order to mitigate architectural drift and 
erosion, and architectural knowledge vaporization.

@@

\item Marlon Pierce: Patching It Up, Pulling It Forward

Pierce discussed how open open source is. Open software needs a diverse, openly governed
community behind it, just as it needs open licensing and a public code
repository. To probe the level of governance within open source projects, he and
his co-authors (Suresh Marru and Chris Mattman)
suggested a contest to encourage individual developers to submit patches and pull
request to projects that are important to them. This simple mechanism will
expose several governance mechanisms, such as how easy it is for independent
developers to communicate with project leadership, how projects accept and
properly license third party contributions, and how projects make decisions such
as granting source tree write access.

@@

\item Samin Ishtiaq: Daemons, Notifications and Sustaining Software
%  Microsoft Research Cambridge

The reproduction and replication of novel results has become a major issue in
computer science, systems biology, and other computational disciplines. These
include both the inability to re-implement novel algorithms and approaches, and
lack of an agreement on how and what to benchmark these algorithms on.
Ishtiaq pointed out these problems and made several suggestions to address them.

\item James Howison: Retract all Bit-Rotten Publications

Howison sought to provoke discussion by proposing that papers whose workflows are
not kept current with the changing software ecosystem should be automatically
retracted. This would create an incentive for authors to keep their
software current and usable, rather than the current situation in which every
potential user has to do this individually. A softer version of the proposal
would identify papers whose software workflow has become bit-rotten and allow
others to keep the code up to date, either adding them as new authors of the
paper or providing credit in some other form for their academic service.

@@

\item Carl Boettiger: rOpenSci: building sustainable software by fostering a
diverse community

Boettiger described how the rOpenSci project has been successful by focusing not just on
building software but also on building a community of researchers who learn and adopt
their approaches to reproducible research and sustainable software practice.
Through outreach, mentoring, workshops, and hackathons, they have not only reached
new users, but also turned users into co-developers of robust software and good practices to
support data science research across a growing set of disciplines.

@@

\item Alice Allen: Find it! Cite it!

>>>>

\begin{comment}
Lighting talks, breakout groups, working groups are all different. Lightning
talks were new in WSSSPE2. They were offered to all paper contributors, and some
other attendees, those who got travel awards. Breakout groups in WSSSPE1 were
just discussions about topics; in WSSSPE2, the breakout groups were supposed to
come up with plans for actions in specific areas. The idea of working groups is
that they will come out of breakout groups as a way for the groups to actually
do the actions over time, not at the workshops.

Email addresses of speakers:
gvwilson@software-carpentry.org, Colin Venters <C.Venters@hud.ac.uk>,
j.spencer@imperial.ac.uk, erinrobinson@esipfed.org, marpierc@iu.edu,
jwpeterson@gmail.com, abani@buffalo.edu, clenhardt@renci.org, dsk@uchicago.edu,
Samin.Ishtiaq@microsoft.com, james@howison.name, s.a.harris@leeds.ac.uk,
marcus.hanwell@kitware.com, thabermann@hdfgroup.org, rdowns@ciesin.columbia.edu,
cboettig@gmail.com, jakob.blomer@cern.ch, editor@ascl.net 
\end{comment}

@@

\item \textbf{Colin Venters: The Nebuchadnezzar Effect: Dreaming of Sustainable Software
through Sustainable Software Architectures.}
Venters proposed that technically sustainable software is a composite, first-class, 
non-functional requirement (NFR) that is at a minimum a measure of a systems 
maintainability and extensibility but may also include other NFR such as efficiency 
(energy and cost), interoperability, portability, reusability, scalability, and usability. To 
achieve technically sustainable software Venters suggested that software 
architectures are fundamental as they are the primary carrier of system qualities 
(NFR), i.e., pre-system understanding, and influence how developers are able to 
understand, analyze, extend, test and maintain a software system, i.e., post-
deployment system understanding. In addition, Venters highlighted that the 
sustainability of software architectures also needs to be addressed to endure 
different types of change and evolution in order to mitigate architectural drift and 
erosion, and architectural knowledge vaporization.

@@

\item \textbf{Marlon Pierce: Patching It Up, Pulling It Forward.}
Pierce discussed how open open source is. Open software needs a diverse, openly governed
community behind it, just as it needs open licensing and a public code
repository. To probe the level of governance within open source projects, he and
his co-authors (Suresh Marru and Chris Mattman)
suggested a contest to encourage individual developers to submit patches and pull
request to projects that are important to them. This simple mechanism will
expose several governance mechanisms, such as how easy it is for independent
developers to communicate with project leadership, how projects accept and
properly license third party contributions, and how projects make decisions such
as granting source tree write access.

@@

\item \textbf{Samin Ishtiaq: Daemons, Notifications and Sustaining Software.}
%  Microsoft Research Cambridge
The reproduction and replication of novel results has become a major issue in
computer science, systems biology, and other computational disciplines. These
include both the inability to re-implement novel algorithms and approaches, and
lack of an agreement on how and what to benchmark these algorithms on.
Ishtiaq from Microsoft Research Cambridge pointed out these problems and made several suggestions to address them.

\item \textbf{James Howison: Retract all Bit-Rotten Publications.}
Howison sought to provoke discussion by proposing that papers whose workflows are
not kept current with the changing software ecosystem should be automatically
retracted. This would create an incentive for authors to keep their
software current and usable, rather than the current situation in which every
potential user has to do this individually. A softer version of the proposal
would identify papers whose software workflow has become bit-rotten and allow
others to keep the code up to date, either adding them as new authors of the
paper or providing credit in some other form for their academic service.

@@

\item \textbf{Carl Boettiger: rOpenSci: Building Sustainable Software by Fostering a
Diverse Community.}
Boettiger described how the rOpenSci project has been successful by focusing not just on
building software but also on building a community of researchers who learn and adopt
their approaches to reproducible research and sustainable software practice.
Through outreach, mentoring, workshops, and hackathons, they have not only reached
new users, but also turned users into co-developers of robust software and good practices to
support data science research across a growing set of disciplines.

@@

\item \textbf{Alice Allen: Find it! Cite it!}
